Standardized crypto-loans on\\ the Cardano blockchain

### 2021-03-24 papers/actus2021/article.tex
Minor edits and tidyings up prior to submission of version for publication.

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Crypto-loans are %unique and 
innovative financial instruments that
allow trustless peer-to-peer %P2P 
lending, and potentially providing a safe and convenient
source of liquidity for cryptocurrency holders. In this paper we
explore a smart contract framework for building standardized crypto-loans
using the Marlowe domain-specific language and with the ACTUS standard at its core.

@@

Smart contracts -- programs that run in a blockchain environment --  can be defined in a variety of ways~\cite{cryptoeprint:2016:1156}. Many such languages are general purpose, and can be used to program any kind of contract it makes sense to run on a blockchain; moreover, they tend to be expressive enough to be Turing complete (in some cases with restrictions on the runtime environment). For example, Plutus, the general-purpose language running on the Cardano blockchain~\cite{utxo-vs-accounts}, is a dialect of Haskell. Another approach is to develop special-purpose or \emph{domain-specific languages} (DSLs) which embody a particular application domain: Marlowe~\cite{marlowe} is a high-level DSL for writing
financial contracts on the Cardano blockchain.

@@

Each Marlowe contract has a finite set of possible execution paths, and so it is possible to analyse the complete behaviour of a contract without running it. Such \emph{static analysis}, based on SMT solving~\cite{smt}, can be used to check properties of a contract; for example, it is possible to check whether a contract will honour all \texttt{Pay} constructs, however it is executed. In the case that a \texttt{Pay} can fail, the analysis gives an example trace showing how that failure can happen.
The language design and static analysis provide assurance that Marlowe contracts are much less likely to ``misbehave'' than contracts written in a general-purpose language like Solidity.

@@

Such loans often represent an investment in a particular venture or industry. As
a somewhat fictional example, one could imagine a cryptocurrency miner
who decided to scale their crypto-farm: a loan (in USD) with variable
interest that directly depends on cryptocurrency prices would be more
attractive for a miner because it would directly correlate with miner's
profits. For example, if the price of the cryptocurrency goes down
in a particular month the borrower would have to pay lower interest, and so would always pay a fixed share of the profits. In a more traditional
setup the interest rate could depend on prices of other commodities: the canonical example would be a power plant taking a loan
with interest depending on electricity prices.

@@

\noindent 
\begin{equation*}
\Delta_{r}=\ident{capfloor}(\ident{driver}*\ident{multiplier}+\ident{spread}-\ident{interestRate}_{t-1})
\end{equation*}

@@

\noindent 
\begin{equation*}
\ident{capfloor}(x)=\ident{max(min}(x,\ident{floor}),\ident{cap})
\end{equation*}

\noindent
The \emph{spread} parameter here loosely represents the difference between
the average prime 
% SJT explain "prime" here.
% it's standard financial terminology: https://www.investopedia.com/terms/p/primerate.asp
rate that the lender expects -- the \emph{benchmark yield} -- and
the rate imposed by the driver - the higher the spread, the higher
the resulting interest rate. The multiplier %effectively 
rescales the interest rate curve 
% SJT used value here rather than plot: is that OK?
% dk14 maybe "curve" would be better
in order to represent the changes to be made converting between different units
of measurement: how many rate percentage points 
% SJT percentage points of what?
% dk14 rate
you would get for a USD-to-kilowatt conversion
and so on.

@@

\noindent 
\begin{equation*}
interestPayment=interestScalingFactor*interestRate*notional
\end{equation*}
\noindent
This scaling is dynamic and loosely adjusts for variance (volatility)
of the asset that the interest rate driver represents.

@@

\subsection{Utilizing polymorphism to abstract over basic operations}

@@

In this section we describe how concrete Marlowe contracts are generate from the terms -- i.e. parameters -- of standard ACTUS contracts; we also describe the implementation of the system, and reflect on the limitations of generating contracts in Marlowe, where contracts have a a predefined lifetime and collection of interactions with contract participants.

@@

\noindent 
\begin{equation*}
\ident{chainlink}(t)=\ident{receiveData}(t)\circ \ident{calculatePayoff(t)}\circ \ident{processPayoff(t)}
\end{equation*}

\noindent 
\begin{equation*}
\ident{contract}(ct)=\ident{collaterals}(ct)\circ \ident{INIT}(ct)\circ\prod_{\footnotesize t\in \ident{SCHED(ct)}}\ident{chainlink(t)}
\end{equation*}

@@

\subsection{ActusLabs}

@@

\newcommand{\rf}{\ident{rf}}
\newcommand{\ct}{\ident{ct}}
\smallskip
\noindent 
\begin{equation*}
\forall \ct.\forall \rf. \; \ident{getCashFlows}(\ident{"haskell"},\ct,\rf)\equiv \ident{getCashFlows}(\ident{"java"},\ct,\rf)
\end{equation*}
\noindent
where \emph{ct} represents contract terms, \emph{rf} is risk factor model, and
\emph{getCashFlows()} returns a set of \emph{(date,payoff)} tuples.

@@

Unlike current mainstream DeFi lending approaches - Marlowe ACTUS-based
implementation tends to rely on trade-matching instead of pooling.
While asset pooling is proven to be superior to order-book based approaches
when it comes to automated market makers, using it for lending has
been shown to be more susceptible to attacks~\cite{flash-loan}.

Moving trade matching off-chain also improves scalability of the protocol
- every loan is a separate contract, thus there is no global state.

There are frameworks, like ISDA Common Domain Model \cite{isda-cdm}, providing a
more precise representation of the business processes involved in institutional trading 
as well as code-generation capabilities. However, due to its structural simplicity, ACTUS 
is much more suited for generating code in a financial domain specific language 
like Marlowe rather than general-purpose one like Java or Haskell.

@@

Marlowe language was explicitly designed as a set of building blocks
for financial contracts that could be combined by anyone familiar
with basic programming. Marlowe ACTUS generators improve on that by
providing a way to automatically combine blocks based on standardized
requirements specified by the user.

On top of that, Marlowe ACTUS provides a toolkit for cross-testing
with original ACTUS spec, a framework for adding new contract types,
cash-flow visualisations and verification tooling.

We are very grateful to Finley and Kegan McIlwaine of the University of Wyoming for their implementation of the LAM contract in ActusLabs, and to colleagues at Quanterall%\footnote{\url{https://quanterall.com/}} 
for their contributions to the Marlowe ACTUS codebase.

>>>>

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Crypto-loans are %unique and 
innovative financial instruments that
allow trustless peer-to-peer %P2P 
lending, and potentially providing a safe and convenient
source of liquidity for cryptocurrency holders. In this paper we
explore a smart contract framework for building standardised crypto-loans
using the Marlowe domain-specific language and the ACTUS standard for financial contracts.

@@

Smart contracts -- programs that run in a blockchain environment --  can be defined in a variety of ways~\cite{cryptoeprint:2016:1156}. Many such approaches are general purpose, and can be used to program any kind of contract it makes sense to run on a blockchain; moreover, they tend to be expressive enough to be Turing complete (in some cases with restrictions on the runtime environment). For example, Plutus, the general-purpose language running on the Cardano blockchain~\cite{utxo-vs-accounts}, is a dialect of Haskell. Another approach is to develop special-purpose or \emph{domain-specific languages} (DSLs) which embody a particular application domain: Marlowe~\cite{marlowe} is a high-level DSL for writing
financial contracts on the Cardano blockchain.

@@

Each Marlowe contract has a finite set of possible execution paths, and so it is possible to analyse the complete behaviour of a contract without running it. Such \emph{static analysis}, based on SMT solving~\cite{smt}, can be used to check properties of a contract; for example, it is possible to check whether a contract will honour all the \texttt{Pay} constructs that it contains, however it is executed. In the case that a \texttt{Pay} can fail, the analysis gives an example trace showing how that failure happens.
The language design and static analysis provide assurance that Marlowe contracts are much less likely to ``misbehave'' than contracts written in a general-purpose language like Solidity.

@@

Such loans often represent an investment in a particular venture or industry. As
a somewhat fictional example, one could imagine a cryptocurrency miner
who decided to scale their crypto-farm: a loan (in USD) with variable
interest that directly depends on cryptocurrency prices would be more
attractive for a miner because it would directly correlate with miner's
profits. For example, if the price of the cryptocurrency goes down
in a particular month the borrower would have to pay lower interest, and so would always pay a fixed share of the profits. In a more traditional
setup the interest rate could depend on prices of other commodities: a canonical example would be a power plant taking a loan
with interest depending on electricity prices.

@@

\smallskip
\noindent 
\begin{equation*}
\Delta_{r}=\ident{capfloor}(\ident{driver}*\ident{multiplier}+\ident{spread}-\ident{interestRate}_{t-1})
\end{equation*}

@@

\smallskip
\noindent 
\begin{equation*}
\ident{capfloor}(x)=\ident{max(min}(x,\ident{floor}),\ident{cap})
\end{equation*}

\noindent
The \emph{spread} parameter here loosely represents the difference between
the average prime 
% SJT explain "prime" here.
% it's standard financial terminology: https://www.investopedia.com/terms/p/primerate.asp
rate that the lender expects -- the \emph{benchmark yield} -- and
the rate imposed by the driver: the higher the spread, the higher
the resulting interest rate. The multiplier %effectively 
rescales the interest rate curve 
% SJT used value here rather than plot: is that OK?
% dk14 maybe "curve" would be better
in order to represent the changes to be made converting between different units
of measurement: how many rate percentage points 
% SJT percentage points of what?
% dk14 rate
you would get for a USD-to-kilowatt conversion
and so on.

@@

\smallskip
\noindent 
\begin{equation*}
interestPayment=interestScalingFactor*interestRate*notional
\end{equation*}
\noindent
This scaling is dynamic and loosely adjusts for variance (volatility)
of the asset that the interest rate driver represents.

@@

\subsection{Utilising polymorphism to abstract over basic operations}

@@

In this section we describe how concrete Marlowe contracts are generated from the terms -- i.e. parameters -- of standard ACTUS contracts. We also describe the implementation of the system, and reflect on the limitations of generating contracts in Marlowe, where contracts have a predefined lifetime and a predefined collection of interactions with contract participants.

@@

\medskip
\noindent 
\begin{equation*}
\ident{chainlink}(t)=\ident{receiveData}(t)\circ \ident{calculatePayoff(t)}\circ \ident{processPayoff(t)}
\end{equation*}

\medskip
\noindent 
\begin{equation*}
\ident{contract}(ct)=\ident{collaterals}(ct)\circ \ident{INIT}(ct)\circ\prod_{\footnotesize t\in \ident{SCHED(ct)}}\ident{chainlink(t)}
\end{equation*}

@@

\subsection{Actus Labs}

@@

\smallskip
\newcommand{\rf}{\ident{rf}}
\newcommand{\ct}{\ident{ct}}
\smallskip
\noindent 
\begin{equation*}
\forall \ct.\forall \rf. \; \ident{getCashFlows}(\ident{"haskell"},\ct,\rf)\equiv \ident{getCashFlows}(\ident{"java"},\ct,\rf)
\end{equation*}
\noindent
where \emph{ct} represents contract terms, \emph{rf} is risk factor model, and
\emph{getCashFlows()} returns a set of \emph{(date,payoff)} tuples.

@@

Unlike current mainstream DeFi lending approaches, our Marlowe ACTUS implementation relies on trade-matching instead of pooling.
While asset pooling is proven to be superior to order-book based approaches
when it comes to automated market makers, using it for lending has
been shown to be more susceptible to attacks~\cite{flash-loan}.
Moving trade matching off-chain also improves scalability of the protocol
- every loan is a separate contract, thus there is no global state.

There are frameworks, like ISDA Common Domain Model \cite{isda-cdm}, providing a
more precise representation of the business processes involved in institutional trading 
as well as code-generation capabilities. However, due to its structural simplicity, ACTUS 
is more suited for generating code in a financial domain specific language 
like Marlowe rather than general-purpose one like Java or Haskell.

@@

The Marlowe language was explicitly designed as a set of building blocks
for financial contracts that could be combined by anyone familiar
with basic programming. The Marlowe ACTUS generators improve on that by
providing a way to automatically combine blocks based on standardised
requirements specified by the user. Marlowe ACTUS also provides a toolkit for cross-testing
against the original ACTUS spec, a framework for adding new contract types,
cash-flow visualisations and verification tooling.

We are very grateful to colleagues at Quanterall and Finley and Kegan McIlwaine of the University of Wyoming for their %implementation of the LAM contract in Actus Labs, and to colleagues at Quanterall %\footnote{\url{https://quanterall.com/}} 
contributions to this project.
